---
title: "Assignment 3 and 4 HealthCare Management Ryerson"
author: "Faizan Khalid Mohsin"
date: "March 6, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 3. Data description and t-tests (10%, split evenly between a. and b.)
a. Use the GE data.
Compare, using t-tests, the ages and BMI of the sample by gender.

```{r}

data_GE = read.csv("GE data W 2018 .csv") # YEs, there is a space in the name. 
str(data_GE)

males_age = subset(data_GE, gender=="Male", age)
females_age = subset(data_GE, gender=="Female", age)

t.test(males_age, females_age)

males_BMI = subset(data_GE, gender=="Male", bmi)
females_BMI = subset(data_GE, gender=="Female", bmi)

t.test(males_BMI, females_BMI)

```

b. Use the California Hospital data.


i. Compare hospital size (number of beds) by type of control (Use groups from Assignment 1)

```{r}
data_California = read.csv("California hospital data W 2018.csv")


```


ii. Compare, using ANOVA, the average discharges, average total revenues and average number of physicians, by hospital size categories (use number of beds for size, 4 categories).

```{r}
fit <- aov(y ~ A, data=data_California)
summary(fit)
```

iii. Compare nurse (regular) and administrative salaries by hospital size categories.

```{r}

subset(data,hosp_size_cat ==asfa , c("nurse_reg", "admin_salary") )

x = c(1.2, 3.2, 4.1, 1)
sd(x)^2
Y = cbind()
manova(Y ~ hosp_size_cat, data = data_California)

```


iv. Compare mortality rates for stroke and AMI by hospital size categories.


```{r }



```

## Assignment 4:


Assignment 4. Regression & correlation (15%)
a. Use the GE data
i.  Make a correlation table using the variables stroke, chd, diabetes, BMI, age, gender and smoking. Note you will need to make dummy variables to do this. 

```{r}

require(corrplot)
require(dplyr)
require(dummies)
library(Hmisc)

data_GE$smoking=as.character(data_GE$smoking)
data_GE$smoking[data_GE$smoking==""] = NA
data_GE$smoking=as.factor(data_GE$smoking)
str(data_GE)
GE = select(data_GE, stroke, chd, diab, bmi, age, gender, smoking )
dput(names(GE))
str(GE)

dat <- dummy.data.frame(GE)
which.dummy(dat)
cor(dat, use="pairwise.complete.obs", method="kendall")
cor(dat, use="pairwise.complete.obs", method="spearman")

# rcorr(dat, type="pearson")
# M = cor(dat)
# corrplot(M)

dput(names(dat))

print(c("strokeNo", "strokeYes", "chdNo", "chdYes", "diabNo", "diabYes", 
"bmi", "age", "genderFemale", "genderMale", "smokingcurrent", 
"smokingever", "smokingformer", "smokingnever", "smokingnot current", 
"smokingNA"), quote = F)

print(c("stroke", "chd", "diab", "bmi", "age", "gender", "smoking"), quote = F)
```

ii.  What factors are associated with the incidence of stroke? Do this with a regression controlling for the different factors.  (use logistic regression – logit)



```{r}

GE = select(data_GE, stroke, chd, diab, bmi, age, gender, smoking, hypertension )

model2 <- glm(stroke ~  chd   +  diab +  bmi  +  age  +   gender + smoking,family=binomial(link='logit'), data=GE)

summary(model2)
# 
# model1 <- glm(stroke ~ chdNo + chdYes+   diabNo    +         diabYes           
#  + bmi             +   age +               genderFemale      
# + genderMale  +       smokingcurrent +    smokingever      
# + smokingformer+      smokingnever  +     smokingnot current
# + smokingNA, family=binomial(link='logit'), data=dat)

```


Iii.  What factors are associated with the incidence of CHD? Do this with a regression controlling for the different factors. What factors are associated with hypertension?
```{r}
model3 <- glm(  chd   ~ stroke + diab +  bmi  +  age  +   gender + smoking,family=binomial(link='logit'), data=GE)
summary(model3)

model4 <- glm( hypertension~  chd   + stroke + diab +  bmi  +  age  +   gender + smoking,family=binomial(link='logit'), data=GE)
summary(model3)

```


b. Use the California Hospital data
i.  What factors are associated with hospital quality – as measured by mortality rates – associated with treatment for AMI and Stroke?

```{r}
select(data_California, contains("mortality"))

mortality = select(data_California, contains("rate"))

```




